---
title: "Genetic matching using R"
subtitle: ""  
author: 
  - "Emma Vestesson"
  - "Paris Pariza"
date: ''
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(gtsummary)
library(skimr)
library(broom)
library(Matching)
library(rgenoud)
library(parallel)
df <- readRDS(here::here('data', "df.rds"))
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)


style_duo_accent(
  primary_color = "#DD0031",
  secondary_color = "#7EBFDA",
  inverse_header_color = "#744284",
)
```

# Prep

Code and slides available on github https://github.com/THF-evaluative-analytics/webinar-genetic-matching

We will be using some custom created data so you will need to download the project if you want to follow along

Install all packages in the workshop_for_participants.Rmd (in the R folder)


---

# About the improvement analytics unit

- an innovative partnership between NHS England and the Health Foundation that provides robust analysis to help the NHS improve care for patients. 

- Our aim is that our analysis helps the local NHS and its partners identify whether implementation of an initiative is having the  desired effect, or needs to change to succeed. 

---
# Agenda

- R and rmarkdown refresher
- 
---
# R refresher

R Studio projects make it straightforward to divide your work into multiple contexts, each with their own working directory, workspace, history, and source documents. 

This means that:

- The .Rprofile file in the project's main directory is sourced by R.
- The current working directory is set to the project directory.
- Previously edited source documents or settings are restored.

---

- tidyverse is collection of R packages based on tidy data principles.
- dplyr package allows simple commands for data manipulation (select(), filter(), mutate(), group_by()).
- pipes (%>%) simplify R code by passing data through functions without nesting. 

---

# Rmarkdown

R Markdown is a file format for making dynamic documents with R and it contains chunks of embedded R code.
- You can knit the file and you can convert the file into an HTML, PDF, or Microsoft Word file.
- Use one hashtag, #, to create a header or two, ##, to create a second level header, and so on.
- To omit the code from the final report, add the argument echo = FALSE inside the {r}. 
- To avoid evaluating a chunk of R code, add the argument eval = FALSE inside the {r}.

---

# Exploratory data analysis

(0) Open the “project.Rproj”, then the file “workshop_participants.Rmd” and load the libraries required

(1) Read the dataset “df.rds” in R
“df” is a completely fake data set that we have generated for this workshop and it is intended to replicate a patient level data where some patients received an intervention aimed at reducing a negative outcome. 


---

# Which variable is **not** in the data set?

a)  `hypertension`

b)  `diabetes`

c)  `stroke`



---

```{r age-histogram, results = "hide"}
median_df <- df %>% 
  group_by(intervention) %>% 
  summarise(age=median(age))

ggplot(df, aes(x=age, ..density..), fill='pink') + geom_histogram(binwidth=1) +  
  geom_vline(data=median_df, aes(xintercept=age),   col='purple', size=1) +  geom_text(data=median_df, aes(x=age, y=0.1, label= paste0('Median:', round(age,digits=0))),                         hjust=-0.5, size=3) + facet_wrap(~intervention) +                           theme_minimal() +                            theme(legend.position = "none")
```


---
# Retrospective matched control studies

Randomised control designs, are the “gold standard” of causal impact evaluation. Often these are impractical, unethical or irrelevant when assessing the impact of complex changes to health delivery.

Observational studies are an alternative approach, where intervention and control groups are as similar as possible. 

One approach to do this is by matching, which can reduce bias in the estimation of the treatment effect. 

Types of matching: 

	~ Nearest neighbour matching: 1-1
	~ Propensity score matching
	~ Genetic Matching
	
---

### The impact of providing enhanced support for care home residents in Rushcliffe (IAU 2017): 
Care home residents receiving enhance support in Rushcliffe matched with individuals with similar characteristics living in care homes in comparable 	areas elsewhere the country.

### Craniectomy in patients with cerebral artery stroke (Fung et al. 2012): 
Patients undergoing craniectomy for cerebral artery stroke were matched to patients undergoing best medical treatment on a range of clinical, 	radiological and surgical characteristics.



---
class: middle

# Genetic matching
---
## Deciding what variables to include to the matching

- Priority should be given to variables predictive of the outcome.

--

- Never include variables that may have been affected by the treatment of interest.

--

- The more variables you include, the longer it will take for matching to run.

--

- There may still be concern regarding unobserved differences between the intervention and control groups.
---
# Other things to consider
 
Match with replacement or not?

--

How many controls for each treated individual? 

--
- Use local or non-local controls?

